---
title: 'Data Analytics 101 Course Project: Car Price Predictor'
author: "Wisdom Agbeve & Daniel Enesi"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: inline
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Brief introduction of the project:**
This project aims to build predictive models for car prices using two datasets: Cars93 and CarPrice_Assignment


**Load both cars data**
```{r}
cars93 <- read.csv("Cars93.csv")
cars_assign = read.csv("CarPrice_Assignment.csv")
```

**Standard deviations for cars_assign**
```{r}
sd(cars_assign$enginesize)
sd(cars_assign$horsepower)
sd(cars_assign$curbweight)
sd(cars_assign$price)
```

**Standard deviations for cars93**
```{r}
sd(cars93$EngineSize)
sd(cars93$Horsepower)
sd(cars93$Weight)
```

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)  # Ensure ggplot2 is loaded
```


**Dataset 1 Summary**
```{r}
summary(cars_assign)
```

**Dataset 2 Summary**
```{r}
summary(cars93)
```

Scatter
```{r}
cars93$Passengers
```

```{r}
cars93$Cylinders
```

```{r}
columns93 <- c("MPG.city", "MPG.highway", "EngineSize", "Horsepower", "RPM", "Rev.per.mile", "Fuel.tank.capacity", "Length", "Wheelbase", "Width", "Rear.seat.room", "Luggage.room", "Weight")

length(columns93)
```

```{r}
ggplot(cars93, aes(x = MPG.city, y = Price)) + 
  geom_point()
```

```{r}
cars93[]
```

```{r}
for (var in columns93)
{
  print(var);
  p <- ggplot(cars93, aes_string(x = var, y = "Price")) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE);
  print(p);
  correlation <- cor(cars93$Price, cars93[[var]]);
  print(correlation)
}
```

```{r}
columns_assign <- c("wheelbase", "carlength", "carwidth", "carheight", "curbweight", "enginesize", "horsepower", "peakrpm", "citympg", "highwaympg", "price")
```

```{r}
for (var in columns_assign)
{
  print(var);
  p <- ggplot(cars_assign, aes_string(x = var, y = "price")) + 
   geom_point() +
  geom_smooth(method = "lm", se = FALSE);
  print(p);
  correlation <- cor(cars_assign$price, cars_assign[[var]]);
  print(correlation)
}
```

**Cars93 linear regression model**
```{r}
model93 <- lm(Price ~ MPG.city + EngineSize + Horsepower + Fuel.tank.capacity + Weight, data = cars93)
summary(model93)
```

**cars_assignment linear regression model**
```{r}
model_assign <- lm(price ~ carwidth + curbweight + enginesize + horsepower + citympg, data = cars_assign)
summary(model_assign)
```

mpgcity+highway horsepower enginesize weight

**cars_assign - Correlation between pairs of variables except price**
```{r}
for (var in c("carwidth", "curbweight", "enginesize", "horsepower", "citympg")) {
  for (var1 in c("carwidth", "curbweight", "enginesize", "horsepower", "citympg")) {
    
    if (var != var1) {
      print(paste("x =", var, ", y =", var1))
      
      p <- ggplot(cars_assign, aes_string(x = var, y = var1)) + 
        geom_point() +
        geom_smooth(method = "lm", se = FALSE)
      print(p)
      
      correlation <- cor(cars_assign[[var1]], cars_assign[[var]])
      print(paste("Correlation between", var, "and", var1, ":", correlation))
    }
  }
}

```

**cars93 - Correlation between pairs of variables except price**
```{r}
for (var in c("MPG.city", "EngineSize", "Horsepower", "Fuel.tank.capacity", "Weight")) {
  for (var1 in c("MPG.city", "EngineSize", "Horsepower", "Fuel.tank.capacity", "Weight")) {
    
    if (var != var1) {
      print(paste("x =", var, ", y =", var1))
      
      p <- ggplot(cars93, aes_string(x = var, y = var1)) + 
        geom_point() +
        geom_smooth(method = "lm", se = FALSE)
      print(p)
      
      correlation <- cor(cars93[[var1]], cars93[[var]])
      print(paste("Correlation between", var, "and", var1, ":", correlation))
    }
  }
}

```

**Linear regression model using common columns: mpgcity+highway, horsepower, enginesize, weight**
```{r}
model_assign2 <- lm(price ~ citympg + highwaympg + curbweight + enginesize + horsepower, data = cars_assign)
summary(model_assign2)
```

```{r}
summary(model_assign2)
```

**The main model**
```{r}
model_assign3 <- lm(price ~ curbweight + enginesize + horsepower, data = cars_assign)
summary(model_assign3)
```

```{r}
model_assign_all <- lm(price ~ wheelbase + carlength + carwidth + carheight + curbweight + enginesize + horsepower + peakrpm + citympg + highwaympg, data = cars_assign)
summary(model_assign_all)
```

```{r}
any(is.na(cars93))
```

```{r}
custom_model <- lm(price ~ enginesize, data = cars_assign)
summary(custom_model)
```

**Data Cleaning:**
**Clean Cars93.csv, use it to validate model.**
```{r}
price_prediction_model <- model_assign3
```

**Get the columns needed to validate from cars 93, three needed columns**
```{r}
car_variables <- cars93[, c("Weight", "EngineSize", "Horsepower")]
```


**Convert EngineSize to cubic centiliters**
```{r}
car_variables$EngineSize <- car_variables$EngineSize * 100 
```

**Change the column names to reflect that of cars_assign**
```{r}
colnames(car_variables) <- c("curbweight", "enginesize", "horsepower")
```

**Check for null rows**
```{r}
anyNA(car_variables)
```
**Check cars 93 for NA**
```{r}
anyNA(cars93)
```


**Predict for first row**
```{r}
first_row = car_variables[1, ]
print(first_row)
first_row <- data.frame(curbweight = 2705, enginesize = 180, horsepower = 140)
```
```{r}
predicted_price <- predict(price_prediction_model, first_row)
print(predicted_price)
```

**Predict for all in cars93**
```{r}
predicted_prices <- predict(price_prediction_model, car_variables)
print(predicted_prices)
```

**Conclusion and Insights:** The project successfully demonstrated the process of predicting car prices using linear regression model on two distinct datasets. By exploring various predictors, such as engine size, horsepower, and weight, we identified key factors influencing car prices. 
The analysis revealed: **Strong predictors:** Variables like horsepower and curb weight consistently showed a significant impact on car prices across both datasets. **Model performance:** Both models achieved reasonable accuracy, with R-squared values indicating a good fit. However, there is room for improvement in capturing price variability. **Dataset differences:** The comparison between the Cars93 and CarPrice_Assignment datasets highlighted differences in market segments and vehicle specifications.
These findings provide valuable insights into how car features affect pricing, which could benefit both consumers and manufacturers. Future work could explore more advanced models or incorporate additional data to improve predictive power and better account for market dynamics.




