---
title: "Car Price Predictor"
author: "Wisdom Agbeve"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Cars93. CarPrice_Assignment.

Scatterplots between price and other car variables for both datasets `cor` to calculate correlation also. To get variables that have strongest correlation with price. Use the top 1-4 variables to make linreg model to predict price. Maybe do models for each of the variables with price.

Homework -\> correlation and scatterplots between variables. horsepower -\> mpg. length -\> engine_size

Load both cars data

```{r}
cars93 <- read.csv("Cars93.csv")
cars_assign = read.csv("CarPrice_Assignment.csv")
```

Scatter

```{r}
cars93$Passengers
```

```{r}
cars93$Cylinders
```

```{r}
columns93 <- c("MPG.city", "MPG.highway", "EngineSize", "Horsepower", "RPM", "Rev.per.mile", "Fuel.tank.capacity", "Length", "Wheelbase", "Width", "Rear.seat.room", "Luggage.room", "Weight")

length(columns93)
```

```{r}
ggplot(cars93, aes(x = MPG.city, y = Price)) + 
  geom_point()
```

```{r}
cars93[]
```

```{r}
for (var in columns93)
{
  print(var);
  p <- ggplot(cars93, aes_string(x = var, y = "Price")) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE);
  print(p);
  correlation <- cor(cars93$Price, cars93[[var]]);
  print(correlation)
}
```

```{r}
columns_assign <- c("wheelbase", "carlength", "carwidth", "carheight", "curbweight", "enginesize", "horsepower", "peakrpm", "citympg", "highwaympg", "price")
```

```{r}
for (var in columns_assign)
{
  print(var);
  p <- ggplot(cars_assign, aes_string(x = var, y = "price")) + 
   geom_point() +
  geom_smooth(method = "lm", se = FALSE);
  print(p);
  correlation <- cor(cars_assign$price, cars_assign[[var]]);
  print(correlation)
}
```

Cars93 linear regression model

```{r}
model93 <- lm(Price ~ MPG.city + EngineSize + Horsepower + Fuel.tank.capacity + Weight, data = cars93)
summary(model93)
```

cars_assignment linear regression model

```{r}
model_assign <- lm(price ~ carwidth + curbweight + enginesize + horsepower + citympg, data = cars_assign)
summary(model_assign)
```

mpgcity+highway horsepower enginesize weight

cars_assign - Correlation between pairs of variables except price

```{r}
for (var in c("carwidth", "curbweight", "enginesize", "horsepower", "citympg")) {
  for (var1 in c("carwidth", "curbweight", "enginesize", "horsepower", "citympg")) {
    
    if (var != var1) {
      print(paste("x =", var, ", y =", var1))
      
      p <- ggplot(cars_assign, aes_string(x = var, y = var1)) + 
        geom_point() +
        geom_smooth(method = "lm", se = FALSE)
      print(p)
      
      correlation <- cor(cars_assign[[var1]], cars_assign[[var]])
      print(paste("Correlation between", var, "and", var1, ":", correlation))
    }
  }
}

```

cars93 - Correlation between pairs of variables except price

```{r}
for (var in c("MPG.city", "EngineSize", "Horsepower", "Fuel.tank.capacity", "Weight")) {
  for (var1 in c("MPG.city", "EngineSize", "Horsepower", "Fuel.tank.capacity", "Weight")) {
    
    if (var != var1) {
      print(paste("x =", var, ", y =", var1))
      
      p <- ggplot(cars93, aes_string(x = var, y = var1)) + 
        geom_point() +
        geom_smooth(method = "lm", se = FALSE)
      print(p)
      
      correlation <- cor(cars93[[var1]], cars93[[var]])
      print(paste("Correlation between", var, "and", var1, ":", correlation))
    }
  }
}

```

Linear regression model using common columns mpgcity+highway horsepower enginesize weight

```{r}
model_assign2 <- lm(price ~ citympg + highwaympg + curbweight + enginesize + horsepower, data = cars_assign)
summary(model_assign2)
```

```{r}
summary(model_assign2)
```

The main model

```{r}
model_assign3 <- lm(price ~ curbweight + enginesize + horsepower, data = cars_assign)
summary(model_assign3)
```

```{r}
model_assign_all <- lm(price ~ wheelbase + carlength + carwidth + carheight + curbweight + enginesize + horsepower + peakrpm + citympg + highwaympg, data = cars_assign)
summary(model_assign_all)
```

```{r}
any(is.na(cars93))
```

```{r}
custom_model <- lm(price ~ enginesize, data = cars_assign)
summary(custom_model)
```

Clean Cars93.csv, use it to validate model.

```{r}
price_prediction_model <- model_assign3
```

Get the columns needed to validate from cars 93, three needed. columns
```{r}
car_variables <- cars93[, c("Weight", "EngineSize", "Horsepower")]
```


Convert EngineSize to cubic centiliters
```{r}
car_variables$EngineSize <- car_variables$EngineSize * 100 
```

Change the column names to reflect that of cars_assign
```{r}
colnames(car_variables) <- c("curbweight", "enginesize", "horsepower")
```

Check for null rows
```{r}
anyNA(car_variables)
```
Check cars 93 for NA
```{r}
anyNA(cars93)
```


Predict for first row
```{r}
first_row = car_variables[1, ]
print(first_row)
first_row <- data.frame(curbweight = 2705, enginesize = 180, horsepower = 140)
```
```{r}
predicted_price <- predict(price_prediction_model, first_row)
print(predicted_price)
```

Predict for all in cars93
```{r}
predicted_prices <- predict(price_prediction_model, car_variables)
print(predicted_prices)
```

